---
title: "Visualising datasets"
output:
  pdf_document: default
  html_notebook: default
---

In this notebook, we present density plots or barcharts for each of the variables in the four datasets being considered.

```{r setup, include = FALSE}
library(tidyverse)
```

The first dataset is MIMIC-III.

```{r mimic, fig.height=8}
mimic <- read_csv("DATA/MIMIC_III/MIMIC_rand_sub.csv",
                  col_types = cols(outcome = col_factor()))

data_long <- mimic %>%
  pivot_longer(!outcome)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_density() + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```
We can also split the data by outcome.

```{r mimic-outcome, fig.height=8}
ggp2 <- ggplot(data_long, aes(x = value, color = outcome)) +
  geom_density() + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

Next up is SYNTHETIC, where the variables are approximately normally distributed.

```{r synthetic, fig.height=12}
synthetic <- read_csv("DATA/SYNTHETIC/SYNTHETIC_1000_SAMPLES_25_FEATURES_25_INFORMATIVE/synthetic_complete.csv",
                       col_types = cols(output = col_factor()))

data_long <- synthetic %>%
  pivot_longer(!output)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_density() + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```
And split by output:

```{r synthetic-output, fig.height=12}
ggp2 <- ggplot(data_long, aes(x = value, color = output)) +
  geom_density() + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

NHSX_COVID19 has some categorical variables.  We do not try to use the original spreadsheet, but rather only the variables we are using in our experiments.

```{r nhsx}
nhs <- read_csv("DATA/NHSX_COVID19/NHSX_COVID19-complete-used-dataset.csv",
                col_types = cols(Death = col_factor()))

# Categorical and ordinal columns
cols_cat <- c(
    "Sex",
    "PMH hypertension",
    "PMH diabetes mellitus type II",
    "PMH CKD",
    "Current ACEi use",
    "Current Angiotension receptor blocker use",
    "Current NSAID used",
    "PMH CVS disease",
    "PMH Lung disease",
    "NEWS2 score on arrival"
)

nhs_numerical <- nhs %>% select(-any_of(cols_cat))

data_long <- nhs_numerical %>%
  pivot_longer(!Death)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_density(na.rm = TRUE) + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

And split by outcome:

```{r nhsx-outcome, fig.height=12}
ggp2 <- ggplot(data_long, aes(x = value, color = Death)) +
  geom_density() + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

We now plot the categorical variables.

```{r nhsx categorical}
nhs_categorical <- nhs %>%
  select(all_of(c(cols_cat, "Death"))) %>%
  mutate(across(everything(), as.factor))

data_long <- nhs_categorical %>%
  pivot_longer(!Death)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_bar(na.rm = TRUE, width = 0.7) + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```
Split by outcome:

```{r nhsx categorical outcome}
ggp2 <- ggplot(data_long, aes(x = value, fill = Death)) +
  geom_bar(na.rm = TRUE, position = "dodge", width = 0.6) + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

We next do the same for the breast cancer dataset.

```{r breast}
breast <- read_csv("DATA/BREAST_CANCER/Breast_cancer_complete_used_onehot.csv",
                   col_types = cols(`Overall Survival Status` = col_factor()))

# Categorical and ordinal columns
cols_cat <- c(
    "Oncotree Code_ACBC",
    "Oncotree Code_BRCA",
    "Oncotree Code_BRCANOS",
    "Oncotree Code_BREAST",
    "Oncotree Code_IDC",
    "Oncotree Code_ILC",
    "Oncotree Code_IMMC",
    "Oncotree Code_MBC",
    "Oncotree Code_MDLC",
    "Overall Patient Receptor Status_HR+/HER2+",
    "Overall Patient Receptor Status_HR+/HER2-",
    "Overall Patient Receptor Status_HR-/HER2+",
    "Overall Patient Receptor Status_Triple Negative",
    "ER Status of the Primary",
    "Metastatic Disease at Last Follow-up",
    "M Stage",
    "Overall Patient HER2 Status",
    "Overall Patient HR Status",
    "PR Status of the Primary",
    "N Stage",
    "Overall Primary Tumor Grade",
    "Stage At Diagnosis",
    "T Stage"
)

breast_numerical <- breast %>% select(-any_of(cols_cat))

data_long <- breast_numerical %>%
  pivot_longer(!`Overall Survival Status`)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_density(na.rm = TRUE) + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```
Splitting by outcome:

```{r breast outcome}
ggp2 <- ggplot(data_long, aes(x = value, color = `Overall Survival Status`)) +
  geom_density() + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

And the categorical variables.

```{r breast categorical, fig.height=8}
breast_categorical <- breast %>%
  select(all_of(c(cols_cat, "Overall Survival Status"))) %>%
  mutate(across(everything(), as.factor))

data_long <- breast_categorical %>%
  pivot_longer(!`Overall Survival Status`)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_bar(na.rm = TRUE, width = 0.7) + 
  facet_wrap(~ name, ncol = 4, scales = "free")
ggp2
```

And split by outcome:

```{r breast categorical outcome, fig.height=12}
ggp2 <- ggplot(data_long, aes(x = value, fill = `Overall Survival Status`)) +
  geom_bar(na.rm = TRUE, position = "dodge", width = 0.6) + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

Finally, we explore the synthetic categorical dataset.

```{r syncat}
syncat <- read_csv("DATA/SYNTHETIC/SYNTHETIC_CATEGORICAL/synthetic_categorical_complete_onehot.csv",
                   col_types = cols(output = col_factor()))

# Categorical and ordinal columns
cols_cat <- c(
    "cat1_A", "cat1_B", "cat1_C", "cat1_D",
    "cat2", "cat3", "cat4", "cat5", "cat6",
    "ord1", "ord2", "ord3", "ord4", "ord5", "ord6"
)

syncat_numerical <- syncat %>% select(-any_of(cols_cat))

data_long <- syncat_numerical %>%
  pivot_longer(!output)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_density(na.rm = TRUE) + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

Splitting by outcome:

```{r syncat outcome}
ggp2 <- ggplot(data_long, aes(x = value, color = output)) +
  geom_density() + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

And the categorical variables.

```{r syncat categorical, fig.height=8}
syncat_categorical <- syncat %>%
  select(all_of(c(cols_cat, "output"))) %>%
  mutate(across(everything(), as.factor))

data_long <- syncat_categorical %>%
  pivot_longer(!output)

ggp2 <- ggplot(data_long, aes(x = value)) +
  geom_bar(na.rm = TRUE, width = 0.7) + 
  facet_wrap(~ name, ncol = 4, scales = "free")
ggp2
```

And split by outcome:

```{r syncat categorical outcome, fig.height=8}
ggp2 <- ggplot(data_long, aes(x = value, fill = output)) +
  geom_bar(na.rm = TRUE, position = "dodge", width = 0.6) + 
  facet_wrap(~ name, ncol = 3, scales = "free")
ggp2
```

